Advanced Reactor Fuel Cycle (ARFC) Molten Salt Reactor Design

### 2016-10-06 MSR-design.tex
I'm fussy with my name on documents and I am happy as last author.
I find the red and green box links created by hyperref very distracting
kophazi gets a year

\documentclass{article}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2ex plus 0.5ex minus 0.2ex}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{textcomp}
\usepackage{placeins}
\usepackage[T1]{fontenc}
\usepackage{gensymb}
\usepackage[utf8]{inputenc}
\usepackage{caption}
\usepackage[export]{adjustbox}
\graphicspath{{./figures/}}

@@

\title{Advanced Reactor Fuel Cycle (ARFC) Molten Salt Reactor Design}
\author{Alexander Lindsay, Katy Huff, Andrei Rykhlevskii} % As you author additions/changes, add
                                      % your name here!!!

@@

A cutaway diagram of the \gls{MSRE} reactor is shown in \cref{fig:MSRE_reactor}. Fuel
enters through the inlet and is distributed evenly along the outside of the
reactor vessel in a small annular region of 1 inch
width. \cite{robertson_msre_1965} The fluid fuel then flows turbulently
downward, cooling the reactor vessel to within 5\textdegree F of the entering
bulk salt temperature (1175 \textdegree F). \cite{robertson_msre_1965} Some \gls{MSRE}
modeling papers (\cite{kophazi_development_????}) refer to this down-flowing
region as the \textit{downcomer}. The fuel salt is collected in the lower plenum
(lower \textit{space}) of the reactor where straightening vanes remove the
salt's rotational motion. The salt then flows upward through approximately 1140
0.4-inch by 1.2-inch fuel channels that are formed by machining grooves in the
sides of the graphite moderator columns \cite{robertson_msre_1965}; these are the
regions where fission occurs. The graphite columns themselves have 2-inch by
2-inch square cross sections. Nominal core volume is 90 ft$^3$, with 70 ft$^3$
occupied by graphite and 20 ft$^3$ occupied by fuel. \cite{robertson_msre_1965}
After flowing through the core channels, the fuel collects in the upper plenum
before exiting through the outlet.

@@

Kophazi developed a 3D time-dependent simulation that was validated against \gls{MSRE}
data. \cite{kophazi_development_????} The model consists of a conventional
neutronics model extended to deal with the precursor drift that is unique to
MSRs and the core's thermal behavior. The thermal model includes the true
geometry of the fuel channels as well as the moderator connecting them. Salt
flow is assumed to be only in the reactor's axial direction, which reduces the
number of non-linear variables that have to be solved for.

@@

with the last term representing the effect of fuel advection. In
\cite{kophazi_development_????}, \cref{eq:neutrons} is solved only in the
reactor core, whereas \cref{eq:precursors} is solved both in the core and in the
\gls{MSRE} external primary circuit. The external loop is modeled as a single pipe
with a volume equal to the total volume of all the external components and
pipelines of the primary circuitry. \cite{kophazi_development_????} It is not
entirely clear how pipe cross sectional areas are handled, which would determine
the transit times of individual pieces of fluid and subsequently the
distribution of delayed neutron production in the external primary circuit. That
is a piece that would likely need to be included in detailed ARFC accident
models.

As pointed out in \cite{kophazi_development_????}, it is not possible to solve
\cref{eq:neutrons,eq:precursors} on the same mesh. \Cref{eq:neutrons} needs to
be solved over the entire reactor cross-section including fuel channels and the
graphite moderator in between, whereas \cref{eq:precursors} is only applicable
within the fuel channels. Kophazi solves this issue by introducing a fuel
fraction parameter, $p_f$, and creating effective fuel flow
channels. \Cref{fig:kophazi_homo} shows how Kophazi transforms the \gls{MSRE} unit
cell into neutronic and precursor elements. The number of effective fuel
channels in the unit cell is determined by the refinement of the neutron grid,
e.g. the \# of channels is equivalent to the number of neutronic elements in a
given \gls{MSRE} unit cell. Consequently, once the domain is discretized (Kophazi uses
the finite volume method), the delayed neutron source term (last term in
\cref{eq:neutrons}) is multiplied by the fuel fraction $p_f$. Homogenization of
the removal, scattering, and fission cross sections allows translation of the
model from the real to effective geometry. Details of the discretization can be
found in \cite{kophazi_development_????}. Unfortunately, neither
\cite{kophazi_development_????} or cited references for the DALTON code,
describe the boundary conditions used for neutrons.

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{kophazi_neutronics_discretization.png}
  \caption{Homogenized model of fuel channels \cite{kophazi_development_????}}
  \label{fig:kophazi_homo}
\end{figure}

@@

Temperature in the fuel is determined by a 1-D convection model in the axial
direction. Moderator temperatures are calculated with a 3-D conduction
model. Coupling between the moderator and fuel temperatures is accomplished
through an axially varying heat transfer coefficient calculated with a Nusselt
correlation; see \cite{kophazi_development_????} for details of the Nusselt
correlation. Thermal expansion of the fuel was not accounted for while
determining the reactor temperature profile; however, expansion is considered
($\kappa = 0.57 mg/cm^3 K$) when determining homogenized neutron
cross-sections. A model for the heat exchanger is included with a heat sink term
proportional to the temperature difference between fuel and coolant salts, with
the constant of proportionality determined from steady-state \gls{MSRE} operational
data. With the exception of the heat exchanger, the thermal code assumes zero
heat flux boundary conditions.

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{kophazi-actual-vs-model-core-geometry.png}
  \caption{Left figure shows snapshot of actual \gls{MSRE} core geometry; to satisfy
    the 1D SCALE/CSAS/XSDRN code, the more regular geometry in the right figure
    is used \cite{kophazi_development_????}}
  \label{fig:real-vs-model}
\end{figure}

Neutron cross sections were collapsed into eight energy groups, distributed
evenly on a logarithmic scale, with additional groups added in resonance and
thermal regions. Control rods were modeled with albedo boundary conditions for
thermal neutrons. More control rod modeling details can be found in
\cite{kophazi_development_????}.

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{3d-neutron-grid.png}
  \caption{Grid for cross section homogenization and neutron cross sections
    \cite{kophazi_development_????}}
  \label{fig:neutron-grid}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{cross-section-of-thermal-grid.png}
  \caption{Cross-section of thermal grid \cite{kophazi_development_????}}
  \label{fig:thermal-grid}
\end{figure}

@@

\begin{enumerate}
\item Test whether \code{gmsh} can generate heterogeneous 3D mesh structure
  like that in the actual \gls{MSRE} geometry.
\item For expected range of temperatures and some defined sequence of
  temperature step sizes, generate table of few-group neutronics cross sections
  with \code{SERPENT}.
  \begin{itemize}
    \item If \code{gmsh}, or that failing \code{cubit}, is capable of generating
      heterogeneous grid, we can generate unique cross regions for each region
    \item If we cannot generate heterogeneous grid, we will have to homogenize
      the cross sections similar to what is done in
      \cite{kophazi_development_????}.
  \end{itemize}
\item Armed with tables for quick interpolation of cross sections, we should be
  able to perform coupled calculations for neutron density and
  temperature.
  \begin{itemize}
    \item As a first pass, the fuel salt flow velocity can be a user input as
      opposed to a non-linear variable that is a function of temperature.
    \item We will need to decide whether neutron and temperature calculations
      should be performed on the same grid, and in transient calculations
      whether those two fields should be updated on the same step-size.
      \begin{itemize}
        \item If not using the same grid and/or step-size, the neutronic and
          temperature calculations will need to be split into their own
          applications and coupling will have to be performed using the MOOSE
          multi-app system.
      \end{itemize}
  \end{itemize}
\end{enumerate}

>>>>

\documentclass{article}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2ex plus 0.5ex minus 0.2ex}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}
\usepackage{textcomp}
\usepackage{placeins}
\usepackage[T1]{fontenc}
\usepackage{gensymb}
\usepackage[utf8]{inputenc}
\usepackage{caption}
\usepackage[export]{adjustbox}
\graphicspath{{./figures/}}

@@

\title{Advanced Reactor Fuel Cycle (ARFC) Molten Salt Reactor Design}
\author{Alexander Lindsay, Andrei Rykhlevskii, Kathryn Huff} % As you author additions/changes, add
                                      % your name here!!!

@@

A cutaway diagram of the \gls{MSRE} reactor is shown in \cref{fig:MSRE_reactor}. Fuel
enters through the inlet and is distributed evenly along the outside of the
reactor vessel in a small annular region of 1 inch
width. \cite{robertson_msre_1965} The fluid fuel then flows turbulently
downward, cooling the reactor vessel to within 5\textdegree F of the entering
bulk salt temperature (1175 \textdegree F). \cite{robertson_msre_1965} Some \gls{MSRE}
modeling papers (\cite{kophazi_development_2009}) refer to this down-flowing
region as the \textit{downcomer}. The fuel salt is collected in the lower plenum
(lower \textit{space}) of the reactor where straightening vanes remove the
salt's rotational motion. The salt then flows upward through approximately 1140
0.4-inch by 1.2-inch fuel channels that are formed by machining grooves in the
sides of the graphite moderator columns \cite{robertson_msre_1965}; these are the
regions where fission occurs. The graphite columns themselves have 2-inch by
2-inch square cross sections. Nominal core volume is 90 ft$^3$, with 70 ft$^3$
occupied by graphite and 20 ft$^3$ occupied by fuel. \cite{robertson_msre_1965}
After flowing through the core channels, the fuel collects in the upper plenum
before exiting through the outlet.

@@

Kophazi developed a 3D time-dependent simulation that was validated against \gls{MSRE}
data. \cite{kophazi_development_2009} The model consists of a conventional
neutronics model extended to deal with the precursor drift that is unique to
MSRs and the core's thermal behavior. The thermal model includes the true
geometry of the fuel channels as well as the moderator connecting them. Salt
flow is assumed to be only in the reactor's axial direction, which reduces the
number of non-linear variables that have to be solved for.

@@

with the last term representing the effect of fuel advection. In
\cite{kophazi_development_2009}, \cref{eq:neutrons} is solved only in the
reactor core, whereas \cref{eq:precursors} is solved both in the core and in the
\gls{MSRE} external primary circuit. The external loop is modeled as a single pipe
with a volume equal to the total volume of all the external components and
pipelines of the primary circuitry. \cite{kophazi_development_2009} It is not
entirely clear how pipe cross sectional areas are handled, which would determine
the transit times of individual pieces of fluid and subsequently the
distribution of delayed neutron production in the external primary circuit. That
is a piece that would likely need to be included in detailed ARFC accident
models.

As pointed out in \cite{kophazi_development_2009}, it is not possible to solve
\cref{eq:neutrons,eq:precursors} on the same mesh. \Cref{eq:neutrons} needs to
be solved over the entire reactor cross-section including fuel channels and the
graphite moderator in between, whereas \cref{eq:precursors} is only applicable
within the fuel channels. Kophazi solves this issue by introducing a fuel
fraction parameter, $p_f$, and creating effective fuel flow
channels. \Cref{fig:kophazi_homo} shows how Kophazi transforms the \gls{MSRE} unit
cell into neutronic and precursor elements. The number of effective fuel
channels in the unit cell is determined by the refinement of the neutron grid,
e.g. the \# of channels is equivalent to the number of neutronic elements in a
given \gls{MSRE} unit cell. Consequently, once the domain is discretized (Kophazi uses
the finite volume method), the delayed neutron source term (last term in
\cref{eq:neutrons}) is multiplied by the fuel fraction $p_f$. Homogenization of
the removal, scattering, and fission cross sections allows translation of the
model from the real to effective geometry. Details of the discretization can be
found in \cite{kophazi_development_2009}. Unfortunately, neither
\cite{kophazi_development_2009} or cited references for the DALTON code,
describe the boundary conditions used for neutrons.

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{kophazi_neutronics_discretization.png}
  \caption{Homogenized model of fuel channels \cite{kophazi_development_2009}}
  \label{fig:kophazi_homo}
\end{figure}

@@

Temperature in the fuel is determined by a 1-D convection model in the axial
direction. Moderator temperatures are calculated with a 3-D conduction
model. Coupling between the moderator and fuel temperatures is accomplished
through an axially varying heat transfer coefficient calculated with a Nusselt
correlation; see \cite{kophazi_development_2009} for details of the Nusselt
correlation. Thermal expansion of the fuel was not accounted for while
determining the reactor temperature profile; however, expansion is considered
($\kappa = 0.57 mg/cm^3 K$) when determining homogenized neutron
cross-sections. A model for the heat exchanger is included with a heat sink term
proportional to the temperature difference between fuel and coolant salts, with
the constant of proportionality determined from steady-state \gls{MSRE} operational
data. With the exception of the heat exchanger, the thermal code assumes zero
heat flux boundary conditions.

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{kophazi-actual-vs-model-core-geometry.png}
  \caption{Left figure shows snapshot of actual \gls{MSRE} core geometry; to satisfy
    the 1D SCALE/CSAS/XSDRN code, the more regular geometry in the right figure
    is used \cite{kophazi_development_2009}}
  \label{fig:real-vs-model}
\end{figure}

Neutron cross sections were collapsed into eight energy groups, distributed
evenly on a logarithmic scale, with additional groups added in resonance and
thermal regions. Control rods were modeled with albedo boundary conditions for
thermal neutrons. More control rod modeling details can be found in
\cite{kophazi_development_2009}.

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{3d-neutron-grid.png}
  \caption{Grid for cross section homogenization and neutron cross sections
    \cite{kophazi_development_2009}}
  \label{fig:neutron-grid}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{cross-section-of-thermal-grid.png}
  \caption{Cross-section of thermal grid \cite{kophazi_development_2009}}
  \label{fig:thermal-grid}
\end{figure}

@@

\begin{enumerate}
\item Test whether \code{gmsh} can generate heterogeneous 3D mesh structure
  like that in the actual \gls{MSRE} geometry.
\item For expected range of temperatures and some defined sequence of
  temperature step sizes, generate table of few-group neutronics cross sections
  with \code{SERPENT}.
  \begin{itemize}
    \item If \code{gmsh}, or that failing \code{cubit}, is capable of generating
      heterogeneous grid, we can generate unique cross regions for each region
    \item If we cannot generate heterogeneous grid, we will have to homogenize
      the cross sections similar to what is done in
      \cite{kophazi_development_2009}.
  \end{itemize}
\item Armed with tables for quick interpolation of cross sections, we should be
  able to perform coupled calculations for neutron density and
  temperature.
  \begin{itemize}
    \item As a first pass, the fuel salt flow velocity can be a user input as
      opposed to a non-linear variable that is a function of temperature.
    \item We will need to decide whether neutron and temperature calculations
      should be performed on the same grid, and in transient calculations
      whether those two fields should be updated on the same step-size.
      \begin{itemize}
        \item If not using the same grid and/or step-size, the neutronic and
          temperature calculations will need to be split into their own
          applications and coupling will have to be performed using the MOOSE
          multi-app system.
      \end{itemize}
  \end{itemize}
\end{enumerate}
